---
title: "p8105_hw6_yz4720"
author: "Jasmine Zhang"
date: "2023-12-02"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

## Problem 1
### Data Import and Tidy

```{r message=FALSE, warning=FALSE}
homicide_df = read_csv("./data/homicide-data.csv", na = c("", "NA", "Unknown")) |> 
  mutate(city_state = str_c(city, state, sep = ", "),
         victim_age = as.numeric(victim_age),
         resolution = case_when(disposition == "Closed without arrest" ~ 0,
                                disposition == "Open/No arrest" ~ 0,
                                disposition == "Closed by arrest" ~ 1)) |> 
  filter(!(city_state %in% c("Tulsa, AL", "Dallas, TX", "Phoenix, AZ", "Kansas City, MO")),
         victim_race %in% c("Black", "White")) |> 
  select(city_state, resolution, victim_age, victim_sex, victim_race)
```

### Baltimore, MD

```{r message=FALSE, warning=FALSE}
baltimore_glm = filter(homicide_df, city_state == "Baltimore, MD") |> 
  glm(resolution ~ victim_age + victim_sex + victim_race, family = binomial(), data = _)
baltimore_glm |> 
  broom::tidy()|> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(OR, OR_CI_lower, OR_CI_upper)
```

### glm all cities

```{r message=FALSE, warning=FALSE}
model_results = 
  homicide_df |> 
  nest(data = -city_state) |> 
  mutate(models = map(data, \(data) glm(resolution ~ victim_age + victim_sex + victim_race, 
                                        family = binomial(), data = data)),
         results = map(models, broom::tidy))  |> 
  select(-models, -data) |> 
  unnest(cols = results)|> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error))|> 
  filter(term == "victim_sexMale") |> 
  select(city_state, OR, OR_CI_lower, OR_CI_upper) 
model_results |>
  slice(1:5) |> 
  knitr::kable(digits = 3)           
```

### plot: OR and CI for all cities 

```{r}
model_results |> 
  mutate(city_state = fct_reorder(city_state, OR)) |> 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = OR_CI_lower, ymax = OR_CI_upper)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

The plot shows OR and CI for each city ordered by the magnitude of OR. Most of the cities have OR less than 1, which means that crimes with male victims have smaller odds of resolution compared to crimes with female victims after adjusting for victim age and race. For cities with CI that doesn't include 1, it suggests a significant difference in resolution rates by sex after adjustment for victim age and race.


## Problem 2

```{r message=FALSE, warning=FALSE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

