---
title: "p8105_hw6_yz4720"
author: "Jasmine Zhang"
date: "2023-12-02"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(modelr)
library(mgcv)
set.seed(1)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

## Problem 1
### Data Import and Tidy

```{r message=FALSE, warning=FALSE}
homicide_df = read_csv("./data/homicide-data.csv", na = c("", "NA", "Unknown")) |> 
  mutate(city_state = str_c(city, state, sep = ", "),
         victim_age = as.numeric(victim_age),
         resolution = case_when(disposition == "Closed without arrest" ~ 0,
                                disposition == "Open/No arrest" ~ 0,
                                disposition == "Closed by arrest" ~ 1)) |> 
  filter(!(city_state %in% c("Tulsa, AL", "Dallas, TX", "Phoenix, AZ", "Kansas City, MO")),
         victim_race %in% c("Black", "White")) |> 
  select(city_state, resolution, victim_age, victim_sex, victim_race)
```

### Baltimore, MD

```{r message=FALSE, warning=FALSE}
baltimore_glm = filter(homicide_df, city_state == "Baltimore, MD") |> 
  glm(resolution ~ victim_age + victim_sex + victim_race, family = binomial(), data = _)
baltimore_glm |> 
  broom::tidy()|> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(OR, OR_CI_lower, OR_CI_upper)
```

### glm all cities

```{r message=FALSE, warning=FALSE}
model_results = 
  homicide_df |> 
  nest(data = -city_state) |> 
  mutate(models = map(data, \(data) glm(resolution ~ victim_age + victim_sex + victim_race, 
                                        family = binomial(), data = data)),
         results = map(models, broom::tidy))  |> 
  select(-models, -data) |> 
  unnest(cols = results)|> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error))|> 
  filter(term == "victim_sexMale") |> 
  select(city_state, OR, OR_CI_lower, OR_CI_upper) 
model_results |>
  slice(1:5) |> 
  knitr::kable(digits = 3)           
```

### plot: OR and CI for all cities 

```{r}
model_results |> 
  mutate(city_state = fct_reorder(city_state, OR)) |> 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = OR_CI_lower, ymax = OR_CI_upper)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

The plot shows OR and CI for each city ordered by the magnitude of OR. Most of the cities have OR less than 1, which means that crimes with male victims have smaller odds of resolution compared to crimes with female victims after adjusting for victim age and race. For cities with CI that doesn't include 1, it suggests a significant difference in resolution rates by sex after adjustment for victim age and race.


## Problem 2

```{r message=FALSE, warning=FALSE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

### bootstrap samples

```{r}
boot_sample = function(df){ 
  sample_frac(df, replace = TRUE)
}

boot_results = tibble(strap_number = 1:5000) |>
  mutate(strap_sample = map(strap_number, \(i) boot_sample(weather_df)),
         models = map(strap_sample, \(df) lm(tmax~ tmin + prcp, data = df)),
         results_r_sq = map(models, broom::glance),
         results_coeff = map(models, broom::tidy))|> 
  select(-models, -strap_sample) 
```

### distribution of estimates r_squared and log(beta1*beta2)

```{r}
boot_results |> 
  unnest(cols1 = results_r_sq,
         cols2 = results_coeff) |>
  select(r.squared) |> 
  ggplot(aes(x = r.squared)) + 
  geom_density() + xlab("estimate of r.squared") 

tmin_coeff = boot_results |>
  unnest(cols = results_coeff) |> 
  filter(term == "tmin") |> 
  pull(estimate)
prcp_coeff = boot_results |>
  unnest(cols = results_coeff) |> 
  filter(term == "prcp") |> 
  pull(estimate)
log_coeff = log(tmin_coeff*prcp_coeff) |> 
  na.omit() |> 
  as.tibble()
log_coeff |> 
  ggplot(aes(x = value)) + 
  geom_density() + xlab("estimate of log(beta_1*beta_2") 
```

The distribution of estimates of r.squared follows roughly a bell-shaped curve, where the value is highly indicating an accurate fit. The distribution of estimates of log(beta1*beta2) is right skewed after omiting the NA values.


### CI for r_squared and log(beta_1*beta_2)

```{r}
r.squared = boot_results |>
  unnest(cols = results_r_sq) |> 
  pull(r.squared) 
quantile(r.squared, c(0.025, 0.975))

log_coeff = log(tmin_coeff*prcp_coeff) |> 
  na.omit()
quantile(log_coeff, c(0.025, 0.975))
```

The 95% CI for `r.squared` is (0.8886697, 0.9406121). The 95% CI for `log(beta*beta2)` is (-8.963404, -4.603361).


## Problem 3
### Data loading

```{r message=FALSE, warning=FALSE}
birthweight_df = read_csv("./data/birthweight.csv", na = c("", "NA", "Unknown")) |> 
  mutate(frace = as.factor(frace),
         mrace = as.factor(mrace),
         babysex = as.factor(babysex)) |> 
  na.omit()
```

### modeling process

```{r}
birthweight_model = birthweight_df |> 
  lm(bwt ~ babysex + bhead + blength + delwt + fincome + frace + gaweeks + malform + 
       menarche + mheight + momage + mrace + parity + pnumlbw + pnumsga + ppbmi +
       ppwt + smoken + wtgain, data = _)
summary(birthweight_model)
```

I started the model building process by first building a full model with all the variables in the dataset. By examing the p value of each variable, I can decide which factors to include in my final proposed model. I selected the variables `babysex` + `bhead` + `blength` + `delwt` + `gaweeks` + `parity` + `smoken` because their p value is less than 0.05, indicating a significant effect on the outcome `bwt`.

### proposed model

```{r}
birthweight_proposed = birthweight_df |> 
  lm(bwt ~ babysex + bhead + blength + delwt + gaweeks + parity + smoken, data = _)
summary(birthweight_proposed)
```

### plot: model residuals vs. fitted values 

```{r}
birthweight_df |> 
  modelr::add_predictions(birthweight_proposed) |> 
  modelr::add_residuals(birthweight_proposed) |> 
  ggplot(aes(x = resid, y = pred)) + geom_point()
```

### compare with two other models with cross-validated prediction error

```{r}
cv_df = birthweight_df |> 
  crossv_mc(n = 100) |>  
  mutate(train = map(train, as_tibble),
         test = map(test, as_tibble))

cv_results = cv_df |> 
  mutate(proposed_mod = map(train, \(df) lm(bwt ~ babysex + bhead + blength + 
                                             delwt + gaweeks + parity + smoken, data = df)),
         maineffect_mod  = map(train, \(df) lm(bwt ~ blength + gaweeks, data = df)),
         interaction_mod  = map(train, \(df) lm(bwt ~ blength + bhead + babysex + 
                                                  babysex*blength + babysex*bhead + babysex*blength*bhead, 
                                                data = df))) |> 
  mutate(rmse_proposed = map2_dbl(proposed_mod, test, \(mod,df) rmse(mod, df)),
         rmse_maineffect = map2_dbl(maineffect_mod, test, \(mod, df) rmse(model = mod, data = df)),
         rmse_interaction = map2_dbl(interaction_mod, test, \(mod, df) rmse(model = mod, data = df)))
```

```{r}
cv_results |> select(starts_with("rmse")) |> 
  pivot_longer(everything(),
               names_to = "model_type", 
               values_to = "rmse", 
               names_prefix = "rmse_") |>
  ggplot(aes(x = model_type, y = rmse)) + geom_violin()
```

